Molecular Gas Kinematics of the Strongly-lensed Quasar Host Galaxy RXS J1131$-$1231

### 2016-11-18 WRITEUP/diff.tex
ready to send to DR

\subsection{Systemic Redshift and Velocity Offset}
%Outflow:
% Not AGN outflows because require different velocity offsets in Balmer and forbidden lines (Zamanov02a)
% line origins:
% \mgii & Hbeta: outer part of the BLR (Goad12a) compared to high-ionization lines (\eg CIV)
% Hbeta can also come from NLR

\citet[]{Sluse07a} report two sets of AGN lines observed in RXJ1131.
The first set of lines is at $z$\ssim\old{0.653}\new{0.654}, including the narrow component of the Balmer lines, the \oiii lines, and the \mgii absorption line; the second set is at $z$\ssim0.658, including the broad component of the Balmer lines and
the \mgii emission line.
Using the CO line center redshift as the \old{systematic} \new{systemic} redshift,
we find that the redshift of the first set is fully consistent with the systemic redshift. This
supports previous claims that [O\,{\scriptsize III}] lines, tracing the narrow line region (NLR),
are good proxies
to the true systemic redshift \citep[e.g.,][]{Vrtilek85a, Nelson00a}.
\old{On the other hand, the second set of lines is blueshifted by $\sim$780\,\kms, indicating that the broad line region (BLR) of the AGN is dynamically offset from the centroid of its host galaxy.}
\new{On the other hand, the second set of lines is redshifted by $\sim$715\,\kms.}

\old{One explanation for the offset may arise from the fact that BLRs are gravitationally bound to
the central AGN, whereas this condition is not necessary for NLRs.
From an SDSS sample of $>$3800 quasars, \mbox{\citet{Richards02a}} find a
median of \mbox{97\kms} and a dispersion of 269\,\mbox{\kms} in the distribution of
the offset between \mbox{[Mg\,{\scriptsize II}] and [O\,{\scriptsize III}]} lines, suggesting
that large offsets are uncommon. Nevertheless,
offsets as large as $\sim$\,500\,\mbox{\kms} have been observed in some cases.}
%% In an SDSS sample of ongoing mergers at $z$\,$<$\,0.21, AGN lines are offset by $\lesssim$410\,km/s
%\citep[\eg][]{Richards02a, Comerford14a}. % between AGN lines (forbidden, Balmer) and stellar absorption

\old{Alternatively, large velocity offsets between emission lines of the
BLR and the NLR may arise in a recoiling black hole (BH),
% recoiling SMBHs as well as subparsec-scale binary SMBHs produce velocity offsets in the broad AGN emission lines and not the narrow AGN emission lines e.g., Gaskell 1983; Bonning et al. 2007; Eracleous et al. 2012; Ju et al. 2013; Shen et al. 2013).
where its BLR is moving at high velocity relative to the bulk of its host galaxy
\mbox{\citep{Madau04a, Bonning07a, Loeb07a}}.
A recoiling black hole is expected
when a pair of uneven mass BH coalesce, during which their orbital energy is being released
as gravitational wave and a
non-zero net angular momentum is being carried away.
Depending on their initial conditions, numerical relativity simulations have shown that
the recoil velocity can reach up to \mbox{$\sim$4000\,\kms} for spinning BHs \mbox{\citep[\eg][]{Campanelli07a}}.
Several sources have been proposed as recoiling BH candidates, such as CID-42 at \mbox{$z$\,=\,0.359 \citep{Civano10a}} and
SDSS 0956+5127 at \mbox{$z$\,$\sim$\,0.714 \citep{Steinhardt12a}}.
One notable system is SDSSJ0927+2943 at \mbox{$z$\,$\sim$\,0.713}, where
a set of redshifted narrow lines and a second set of blueshifted broad AGN lines
have been observed \mbox{\citep{Komossa08a}},  similar to the case for RXJ1131.
However, this scenario was recently refuted by \mbox{\citet{Decarli14a}}, who confirm that the
BLR lines in SDSSJ0927+2943 are consistent with the CO systemic redshift.
This is in contrast with RXJ1131, where our CO observations confirm the blueshifted BLR.
Since this model requires a merged BH, it would imply that RXJ1131 might
have been in a previous merger, which would also explain
the highly spinning BH in RX1131 \mbox{\citep[$a$\ssim0.9;][]{Reis14a}}.}

% REFEREE: offset in BLR and NLR in RXJ1131 has been known before, not new for this paper
%One explanation for the offset may arise from the fact that BLRs are gravitationally bound to
%the central AGN, whereas this condition is not necessary for NLRs.
%From an SDSS sample of $>$3800 quasars, \citet{Richards02a} find a
%median of 97\kms and a dispersion of 269\,\kms in the distribution of
%the offset between [Mg\,{\scriptsize II}] and [O\,{\scriptsize III}] lines, suggesting
%that large offsets are uncommon. Nevertheless,
%offsets as large as $\sim$\,500\,\kms have been observed in some cases.
%%% In an SDSS sample of ongoing mergers at $z$\,$<$\,0.21, AGN lines are offset by $\lesssim$410\,km/s
%%\citep[\eg][]{Richards02a, Comerford14a}. % between AGN lines (forbidden, Balmer) and stellar absorption

%Alternatively, large velocity offsets between emission lines of the
%BLR and the NLR may arise in a recoiling black hole (BH),
%% recoiling SMBHs as well as subparsec-scale binary SMBHs produce velocity offsets in the broad AGN emission lines and not the narrow AGN emission lines (e.g., Gaskell 1983; Bonning et al. 2007; Eracleous et al. 2012; Ju et al. 2013; Shen et al. 2013).
%where its BLR is moving at high velocity relative to the bulk of its host galaxy
%\citep{Madau04a, Bonning07a, Loeb07a}.
%A recoiling black hole is expected
%when a pair of uneven mass BH coalesce, during which their orbital energy is being released
%as gravitational wave and a
%non-zero net angular momentum is being carried away.
%Depending on their initial conditions, numerical relativity simulations have shown that
%the recoil velocity can reach up to $\sim$4000\,\kms for spinning BHs \citep[\eg][]{Campanelli07a}.
%Several sources have been proposed as recoiling BH candidates, such as CID-42 at $z$\,=\,0.359 \citep{Civano10a} and
%SDSS 0956+5127 at $z$\,$\sim$\,0.714 \citep{Steinhardt12a}.
%Since this model requires a merged BH, it would imply that RXJ1131 might
%have been in a previous merger, which would also explain
%the highly spinning BH in RX1131 \citep[$a$\ssim0.9;][]{Reis14a}.

@@

Using the CO redshift as systemic redshift for the host galaxy of RXJ1131, we find
that the redshift inferred from NLR lines reported in previous studies are consistent with this systemic redshift
but the optical BLR lines are redshifted by \old{$\sim$780\,\kms}\new{$\sim$715\,\kms}.
\old{This implies that the AGN is offset from the dynamical center of its host galaxy.
One explanation could be that the
BLRs are gravitationally bound to the central AGN, whereas this condition is not necessary for NLRs.
Since the velocity offset between emission lines of the BLR and the NLR
is typically \mbox{$\sim$100\pmm270\,\kms},
we thus raise an alternative explanation of a recoiling black hole in RXJ1131, which
would also explain the high spin parameter reported by \mbox{\citet{Reis14a}}.}

>>>>

\subsection{\texorpdfstring{\st{Velocity Offset and a Recoiling Black Hole}}{ } Systemic Redshift and Velocity Offset}
%Outflow:
% Not AGN outflows because require different velocity offsets in Balmer and forbidden lines (Zamanov02a)
% line origins:
% \mgii & Hbeta: outer part of the BLR (Goad12a) compared to high-ionization lines (\eg CIV)
% Hbeta can also come from NLR

\citet[]{Sluse07a} report two sets of AGN lines observed in RXJ1131.
The first set of lines is at $z$\ssim\old{0.653}\new{0.654}, including the narrow component of the Balmer lines, the \oiii lines, and the \mgii absorption line; the second set is at $z$\ssim0.658, including the broad component of the Balmer lines and
the \mgii emission line.
Using the CO line center redshift as the systemic redshift,
we find that the redshift of the first set is fully consistent with the \old{systematic} \new{systemic} redshift. This
supports previous claims that [O\,{\scriptsize III}] lines, tracing the narrow line region (NLR),
are good proxies
to the true systemic redshift \citep[e.g.,][]{Vrtilek85a, Nelson00a}.
\old{On the other hand, the second set of lines is blueshifted by $\sim$780\,\kms, indicating that the broad line region (BLR) of the AGN is dynamically offset from the centroid of its host galaxy.}
\new{On the other hand, the second set of lines is redshifted by $\sim$715\,\kms.}
\old{One explanation for the offset may arise from the fact that BLRs are gravitationally bound to
the central AGN, whereas this condition is not necessary for NLRs.
From an SDSS sample of $>$3800 quasars, \mbox{\citet{Richards02a}} find a}
\new{Offset between BLR and NLR lines have been reported in literature, e.g., 
\citet{Richards02a} find a}
 median of 97\kms and a dispersion of \mbox{269\,\kms} in the distribution of
\old{the offset between [Mg\,{\scriptsize II}] and [O\,{\scriptsize III}] lines, suggesting
that large offsets are uncommon. Nevertheless,
offsets as large as \mbox{$\sim$\,500\,\kms} have been observed in some cases.}
\new{the offset between [Mg\,{\scriptsize II}] and [O\,{\scriptsize III}] lines in an SDSS sample of $>$3800 quasars, 
and \mbox{\citet{Bonning07a}} find a mean offset of \mbox{100\,\kms} with a dispersion of \mbox{212\,\kms} %FWHM = 500
between the broad component of
\mbox{H$\beta$ and [O\,{\scriptsize III}]} lines in a sample of $\sim$2600 quasars at 0.1$<$$z$$<$0.81,
illustrating that large offsets (\mbox{$\Delta v$$\gtrsim$\,500\,\kms}) are uncommon but have been observed in some cases.}

\new{Common interpretations invoked to explain offsets between BLR and systemic redshifts include
e.g., outflow/inflow of gas in the BLR, rotation of the accretion disk, viewing angle towards the accretion disk, 
inclination effect, in which clumps in the accretion disk may block emission towards the far side of the disk, and thus the observed BLR emission is only
originating from
the redshifted inflowing gas component in the near side of the accretion disk of RXJ1131.
In the latter, \citet{Sluse07a} report a covering factor of 20\% based on \mgii absorption line at $z$\eq0.654.
Additionally, differentially lensing of the compact BLR emission by microlensing in RXJ1131\citep[\eg][]{Sluse07a, Sluse12a} 
may have boosted the redshifted component more than the blueshifted component,
thereby biasing the BLR line centroids towards longer wavelengths.}
{\bf DR: is it because of the large linewidth (one component of \mgii is $>$20000\kms, and broad comp. of H$\beta$ is $\sim$2000$-$4000\kms) that the referee thinks it's more likely to be inflow and therefore doesn't like the recoil black hole?
Sluse07:
"Another remarkable result is the presence in our spectra of a \mgii absorption doublet blueshifted at $z$\eq0.654. 
According to its velocity shift ($\Delta v$\ssim$-$660\,\kms). 
These lines disappear from the spectrum ratios A/B and C/B. This indicates that the absorbed flux is proportional to the flux coming from the continuum+BLR. This implies that the region at the origin of the absorption must cover both the continuum and the BLRs and that, within the uncertainties, their depths are identical in the spectra of the three images A, B, and C."
ok.. so is it because of this broad absorption component of \mgii that is at the same $z$ as the NLR lines, and therefore implying that the BLR and 
the AGN is not really spatially and kinematically offset, 
but the offset is truly just a manifestation of a rotating disk + other kinematics (+ absorption of continuum by BLR moving towards us, redshfiting the line centroid, + inflow of material + clumps blocking the far side inflow?)
}

\old{Alternatively, large velocity offsets between emission lines of the
BLR and the NLR may arise in a recoiling black hole (BH),}
\new{Alternatively, the observed offset 
may be explained by a more extreme scenario of a recoiling black hole (BH),}
where its BLR is moving at high velocity relative to the bulk of its host galaxy
\mbox{\citep{Madau04a, Bonning07a, Loeb07a}}.
A recoiling black hole is expected
when a pair of uneven mass BH coalesce, during which their orbital energy is being released
as gravitational wave and a
non-zero net angular momentum is being carried away.
Depending on their initial conditions \new{(e.g., black hole mass ratio, spin-orbit orientation, spin magnitude)}, numerical relativity simulations have shown that
the recoil velocity can reach up to \mbox{$\sim$4000\,\kms} for spinning BHs\old{\mbox{\citep[\eg][]{Campanelli07a}}}\new{,with a typical offset recoil velocity of $v_{\rm kick}$\ssim100\,$-$\,500\,\kms 
\mbox{\citep[\eg][]{Libeskind06a,Campanelli07a}}}.
Several sources have been proposed as recoiling BH candidates, such as CID-42 at \mbox{$z$\,=\,0.359 \citep{Civano10a}} and
SDSS 0956+5127 at \mbox{$z$\,$\sim$\,0.714 \citep{Steinhardt12a}}\old{. One notable system is}\new{, and} SDSSJ0927+2943 at \mbox{$z$\,$\sim$\,0.713}, where
a set of redshifted narrow lines and a second set of blueshifted broad AGN lines
have been observed \mbox{\citep{Komossa08a}}\old{, similar to the case for RXJ1131}.
However, this scenario was recently refuted by \mbox{\citet{Decarli14a}}, who confirm that the
BLR lines in SDSSJ0927+2943 are consistent with the CO systemic redshift.
This is in contrast with RXJ1131, where our CO observations confirm the \old{blueshifted}\new{redshifted} BLR.
\old{Since this model requires a merged BH, it would imply that RXJ1131 might
have been in a previous merger, which would also} 
\new{Since this scenario requires a coalesced BH, it would imply that RXJ1131 
is a product of previous merger, which is not implausible and 
might also} explain the highly spinning BH in RX1131 \citep[$a$\ssim0.9;][]{Reis14a}.

\new{The available data do not provide direct evidence favoring any of the above scenarios, 
we therefore cannot rule out any of the above plausible scenarios; $\lesssim$1-kpc
resolution observations probing the morphology, structures and kinematics of the (ionized) gas and 
stellar (e.g., via H$\alpha$ kinematics measurements) would be useful to distinguish between these scenarios.}

@@

Using the CO redshift as systemic redshift for the host galaxy of RXJ1131, we find
that the redshift inferred from NLR lines reported in previous studies are consistent with this systemic redshift
but the optical BLR lines are redshifted by \old{$\sim$780\,\kms}\new{$\sim$715\,\kms}.{\bf MAY ADD STUFF ABOUT RECOIL BH..}
\old{This implies that the AGN is offset from the dynamical center of its host galaxy.
One explanation could be that the
BLRs are gravitationally bound to the central AGN, whereas this condition is not necessary for NLRs.
Since the velocity offset between emission lines of the BLR and the NLR
is typically \mbox{$\sim$100\pmm270\,\kms},
we thus raise an alternative explanation of a recoiling black hole in RXJ1131, which
would also explain the high spin parameter reported by \mbox{\citet{Reis14a}}.}

### 2016-11-18 WRITEUP/ms.tex
update conclusion about recoil, put place holder to remind self to update
ready to send to DR

\citet[]{Sluse07a} report two sets of AGN lines observed in RXJ1131.
The first set of lines is at $z$\ssim0.654, including the narrow component of the Balmer lines, the \oiii lines, and the \mgii absorption line; the second set is at $z$\ssim0.658, including the broad component of the Balmer lines and the \mgii emission line.
Using the CO line center redshift as the systemic redshift,
we find that the redshift of the first set is fully consistent with the systemic redshift. This
supports previous claims that [O\,{\scriptsize III}] lines, tracing the narrow line region (NLR),
are good proxies
to the true systemic redshift \citep[e.g.,][]{Vrtilek85a, Nelson00a}.
On the other hand, the second set of lines is redshifted by $\sim$715\,\kms.
Offset between BLR and NLR lines have been reported in literature, e.g., 
\citet{Richards02a} find a
median of 97\kms and a dispersion of 269\,\kms in the distribution of
the offset between [Mg\,{\scriptsize II}] and [O\,{\scriptsize III}] lines in an SDSS sample of $>$3800 quasars, 
and \citet{Bonning07a} find a mean offset of 100\,\kms with a dispersion of 212\,\kms %FWHM = 500
between the broad component of
H$\beta$ and \oiii lines in a sample of $\sim$2600 quasars at 0.1$<$$z$$<$0.81,
illustrating that large offsets ($\Delta v$$\gtrsim$\,500\,\kms) are uncommon but have been observed in some cases.

@@

Using the CO redshift as systemic redshift for the host galaxy of RXJ1131, we find
that the redshift inferred from NLR lines reported in previous studies are consistent with this systemic redshift
but the optical BLR lines are redshifted by $\sim$715\,\kms.

>>>>

\citet[]{Sluse07a} report two sets of AGN lines observed in RXJ1131.
The first set of lines is at $z$\ssim0.654, including the narrow component of the Balmer lines, the \oiii lines, and the \mgii absorption line; the second set is at $z$\ssim0.658, including the broad component of the Balmer lines and the \mgii emission line.
Using the CO line center redshift as the systemic redshift,
we find that the redshift of the first set is fully consistent with the systemic redshift. This
supports previous claims that [O\,{\scriptsize III}] lines, tracing the narrow line region (NLR),
are good proxies
to the true systemic redshift \citep[e.g.,][]{Vrtilek85a, Nelson00a}.
On the other hand, the second set of lines is redshifted by $\sim$715\,\kms.
Offset between BLR and NLR lines have been reported in literature, e.g., 
\citet{Richards02a} find a
median of 97\kms and a dispersion of 269\,\kms in the distribution of
the offset between [Mg\,{\scriptsize II}] and [O\,{\scriptsize III}] lines in an SDSS sample of $>$3800 quasars, 
and \citet{Bonning07a} find a mean offset of 100\,\kms with a dispersion of 212\,\kms %FWHM = 500
between the broad component of
H$\beta$ and [O\,{\scriptsize III}] lines in a sample of $\sim$2600 quasars at 0.1$<$$z$$<$0.81,
illustrating that large offsets ($\Delta v$$\gtrsim$\,500\,\kms) are uncommon but have been observed in some cases.

@@

Using the CO redshift as systemic redshift for the host galaxy of RXJ1131, we find
that the redshift inferred from NLR lines reported in previous studies are consistent with this systemic redshift
but the optical BLR lines are redshifted by $\sim$715\,\kms. {\bf MAY ADD STUFF ABOUT RECOIL BH..}
